The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-06-07 Fsing/FThresholdsPaper/FThresholds.tex
Minor addition to SpecialFThresholds documentation
Minor changes

The \emph{FrobeniusThresholds} package contains methods to
 test candidate values for an $F$-pure threshold, and even $F$-jumping numbers, even in some singular rings.
Consider a  $\mathbb{Q}$-Gorenstein ring $R$ of characteristic $p>0$, whose index not divisible by $p$.
Given a rational number $t$ and an element $f$ of $R$, the command {\tt isFPT(t,f)} outputs {\tt true} if $t$ is the $F$-pure threshold of $f$, and {\tt false} otherwise.
Moreover, the function {\tt compareFPT} provides further information about the relationship between $t$ and the $F$-pure threshold of $f$; namely,
{\tt compareFPT(t, f)} returns {\tt-1} when $t < \fpt(f)$, {\tt 0} when $t = \fpt(f)$ and {\tt 1} when $t > \fpt(f)$.
Setting the option {\tt IsLocal} to {\tt true} tells the function to consider the $F$-pure threshold at the origin.
The following example illustrates this functionality.

@@

The core function in this package is the {\tt fpt} function.  Throughout this section, let $f$ be a polynomial with coefficients in a finite field of characteristic $p$. When passed the polynomial $f$, the function {\tt fpt} attempts to find the exact value for the $F$-pure threshold of $f$ at the origin, and returns that value, if possible.  Otherwise, it returns lower and upper bounds for the $F$-pure threshold, as demonstrated below.

@@

If the option {\tt UseSpecialAlgorithms} is set to {\tt true} (the default value), then {\tt fpt} first checks whether $f$ is a monomial, a diagonal polynomial, a binomial, a form in two variables, or a product of factors in simple normal crossing, in that order.  The monomials and the simple normal crossing cases are computed as the reciprocal of maximal exponent of the factors.
Outside of that simple case, algorithms of Hern\'andez \cite{HernandezFInvariantsOfDiagonalHyp, HernandezFPureThresholdOfBinomial}, or Hern\'andez and Teixeira \cite{HernandezTeixeiraFThresholdFunctions},  are executed to compute the $F$-pure threshold (cf. \cite{ShibutaTakagiLCThresholds}).

@@

i4 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

      94
o4 = ---
     625

i5 : fpt(x^2*y^6*z^10 + x^10*y^5*z^3) -- a binomial

      997
o5 = ----
     6250

o5 : QQ

i6 : ZZ/5[x,y];

i7 : fpt(x^2*y^6*(x + y)^9*(x + 3*y)^10) -- a form in two variables

      5787
o7 = -----
     78125

o7 : QQ
\end{MyVerbatim}
}
\medspace

@@

i8 : L = {x, y, x + y, x + 3*y};

i9 : m = {2, 6, 9, 10};

i10 : fpt(L, m) == o7

If {\tt UseSpecialAlgorithms} is set to {\tt true} and $f$ does not fall into any of the three forementioned cases, then the function {\tt fpt} next calls {\tt isSimpleNormalCrossing(f)} to check whether the polynomial $f$ is a product of factors that are in simple normal crossing.  If this command returns {\tt true}, the $F$-pure threshold can easily be computed, and {\tt fpt} returns this value.  For more detail on the function {\tt isSimpleNormalCrossing} and on computing the $F$-pure threshold of a polynomial with simple normal crossings, please see the end of this section.

>>>>

The \emph{FrobeniusThresholds} package contains methods to
 test candidate values for an $F$-pure threshold, and even $F$-jumping numbers, even in some singular rings.

Consider a  $\mathbb{Q}$-Gorenstein ring $R$ of characteristic $p>0$, whose index not divisible by $p$.
Given a parameter $t \in \mathbb{Q}$ and an element $f$ of $R$, the command {\tt isFPT(t,f)} outputs {\tt true} if $t$ is the $F$-pure threshold of $f$, and {\tt false} otherwise.
Moreover, the function {\tt compareFPT} provides further information about the relationship between $t$ and the $F$-pure threshold of $f$; namely,
{\tt compareFPT(t, f)} returns {\tt-1} , {\tt 0}, {\tt 1} when $t$ is less than, equal to, or greater than, respectively, the $F$-pure threshold of $f$.  Setting the option {\tt IsLocal} to {\tt true} tells the function to consider the $F$-pure threshold at the origin.  The following example illustrates this functionality.

@@

The core function in the package \emph{FrobeniusThresholds} is called {\tt fpt}.  Throughout this section, let $f$ be a polynomial with coefficients in a finite field of characteristic $p$. When passed the polynomial $f$, the function {\tt fpt} attempts to find the exact value for the $F$-pure threshold of $f$ at the origin, and returns that value, if possible.  Otherwise, it returns lower and upper bounds for the $F$-pure threshold, as demonstrated below.

@@

\subsection*{The option {\tt UseSpecialAlgorithms} for the {\tt fpt} function}

The {\tt fpt} function has the option {\tt UseSpecialAlgorithms}.  When
set to {\tt true} (the default value), {\tt fpt} first checks whether $f$ is a monomial, a diagonal polynomial, a binomial, a form in two variables, or a product of factors in simple normal crossing, in that order.

If $f = x_1^{a_n}\cdots x_n^{a_n}$, then its $F$-pure threshold is simply $\min \{ 1/a_1, \ldots, 1/a_n \}$.  When $f$ is a diagonal polynomial, a binomial, or a form in two variables, 
algorithms of Hern\'andez \cite{HernandezFInvariantsOfDiagonalHyp, HernandezFPureThresholdOfBinomial}, or Hern\'andez and Teixeira \cite{HernandezTeixeiraFThresholdFunctions},  are executed to compute the $F$-pure threshold.  We note that algorithms for computing $F$-pure thresholds of binomials in certain special characteristics appear in \cite{ShibutaTakagiLCThresholds}.

@@

i4 : fpt(x^3*y^4*z^5) -- a monomial

      1
o4 = -
      5

i5 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

      94
o5 = ---
     625

i6 : fpt(x^2*y^6*z^10 + x^10*y^5*z^3) -- a binomial

      997
o6 = ----
     6250

o6 : QQ

i7 : ZZ/5[x,y];

i8 : fpt(x^2*y^6*(x + y)^9*(x + 3*y)^10) -- a form in two variables

      5787
o8 = -----
     78125

o8 : QQ
\end{MyVerbatim}
}
\medspace

@@

i9 : L = {x, y, x + y, x + 3*y};

i10 : m = {2, 6, 9, 10};

i11 : fpt(L, m) == o7

o11 = true
\end{MyVerbatim}
}
\medspace

If {\tt UseSpecialAlgorithms} is set to {\tt true} and $f$ does not fall into any of the four forementioned cases, then the function {\tt fpt} next calls {\tt isSimpleNormalCrossing(f)} to check whether the polynomial $f$ is a product of factors that are in simple normal crossing.  If this command returns {\tt true}, the $F$-pure threshold can easily be computed, and {\tt fpt} returns this value.  For more detail on the function {\tt isSimpleNormalCrossing} and on computing the $F$-pure threshold of a polynomial with simple normal crossings, please see the end of this section.

\subsection*{Searching for the $F$-pure threshold when no special algorithm applies}
